"PA1_template.Rmd1"
===============================

###1. Loading and Pre-preparing data

Following code loads the data :

```{r, echo=TRUE}
setwd("D:/RWorking Directory/RepData_PeerAssessment1")
data <- read.csv("activity.csv")
```

And transforms it into a suitable format:
```{r, echo=TRUE}
steps_sum <- tapply(data$steps, data$date, sum)
steps_sum <- as.data.frame(steps_sum)
steps_sum$steps_sum <- as.numeric(steps_sum$steps_sum)
```

###2. What is mean total number of steps taken per day?

Following chunk of code creates a histogram of total numbers of steps taken per day:

```{r, echo=TRUE}
hist(steps_sum$steps_sum, xlab = "Total number of steps made per day", main= "Histogram of total steps made per day", cex.main = 1.2, cex.lab=0.8, font.lab = 3)
```

Following chunk of code counts the mean and median of the sum of steps taken each day, i.e: median of sum of steps taken on day 1, day 2... etc. and similarly with mean. NA values were ignored.

```{r, echo=TRUE}

mean(steps_sum$steps_sum, na.rm=TRUE)
median(steps_sum$steps_sum, na.rm=TRUE)
```

###3. What is the average daily activity pattern?

Creating a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

```{r, echo=TRUE}
steps_averaged <- tapply(data$steps, data$interval, mean, na.rm=TRUE)
steps_averaged <- as.data.frame(steps_averaged)
steps_averaged$steps_averaged <- as.numeric(steps_averaged$steps_averaged)

plot(steps_averaged$steps_averaged, type="l", main="Steps averaged across all days based on intervals", xlab="Intervals", ylab="Steps Averaged", cex.main = 1.2, cex.lab=0.8, font.lab = 3)

```

Finding interval corresponding to the highest steps average:

```{r, echo=TRUE}
steps_averaged$interval <- rownames(steps_averaged)
steps_averaged[which.max(steps_averaged$steps_averaged),]
```

###4. Imputing missing values

How many missing values are there in the data set (number of rows)
```{r, echo=TRUE}
nrow(data[!complete.cases(data),])
```

Exchanging missing values with the averaged values for appropriate intervals and creating a new data set with the missing values exchanged:

```{r}
nd <- merge(data, steps_averaged, by="interval")
nd <- nd[order(nd$date),]

nd$steps[is.na(nd$steps)] <- nd$steps_averaged[is.na(nd$steps)]

```

Creating a histogram of the total number of steps taken per day:

```{r}
steps_sum2 <- tapply(nd$steps, nd$date, sum)
steps_sum2 <- as.data.frame(steps_sum2)
steps_sum2$steps_sum2 <- as.numeric(steps_sum2$steps_sum2)

hist(steps_sum2$steps_sum2, xlab = "Number of steps made per day", main= "Histogram of total steps made per day (NAs exluded)", cex.main = 1.2, cex.lab=0.8, font.lab=3)
```

Comparing two outcomes: with NAs and w/o NAs based on mean and median

```{r}
mean_na <- mean(steps_sum$steps_sum, na.rm=TRUE)
median_na <- median(steps_sum$steps_sum, na.rm=TRUE)
mean_nar <- mean(steps_sum2$steps_sum2, na.rm=TRUE)
median_nar <- median(steps_sum2$steps_sum2, na.rm=TRUE)
comp_table <- matrix(c(mean_na, median_na, mean_nar, median_nar), ncol=2)
rownames(comp_table) <- c("Mean", "Median")
colnames(comp_table) <- c("With NAs", "W/o NAs")
comp_table
```

###5. Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable with two level information on the dates weekday or weekend:

```{r}

Sys.setlocale("LC_TIME", "English")

nd$day <- as.Date(nd$date)
nd$day <- weekdays(nd$day, abbreviate=FALSE)

nd$day[nd$day %in% c("Saturday", "Sunday")] <- "weekend"
nd$day[nd$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")] <- "weekday"
```

Prepare the data (assign correctly "weekday" and "weekend" values) and create a panel plot for comparison of results:

```{r}

weekdays <- nd[nd$day=="weekday",] 
weekends <- nd[nd$day=="weekend",]
wd <- as.data.frame(tapply(weekdays$steps, weekdays$interval, mean, na.rm=TRUE))
we <- as.data.frame(tapply(weekends$steps, weekends$interval, mean, na.rm=TRUE))

wd$intervals <- rownames(wd)
we$intervals <- rownames(we)

colnames(wd) <- c("steps", "intervals")
colnames(we) <- c("steps", "intervals")

par(mfrow=c(2,1))

plot(we$steps, type="l", main = "Weekends", cex.lab=0.8, cex.axis=0.5, font.lab=3, cex.main=1, xlab="Intervals", ylab="Number of steps", col="blue")
plot(wd$steps, type="l", main = "Weekdays", cex.lab=0.8, cex.axis=0.5, font.lab=3, cex.main=1, xlab="Intervals", ylab="Number of steps", col="blue")


```


